---
title: "World App Pitch"
author: "Shawvin"
date: "9/24/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This is an R Shiny app presentation. The app use the world map data from ggplot2 package to build a 3D model of Earth. It allow use to choose the color and rotate the earth model in three dimension

## Choose the Palette

There are five available palette to choose from. The slide will show the name and display of the palette.
```{r,echo = FALSE}
suppressPackageStartupMessages(library(RColorBrewer))
par(mfrow=c(3,2))
display.brewer.pal(11,"PiYG")
display.brewer.pal(11,"PuOr")
display.brewer.pal(11,"RdBu")
display.brewer.pal(11,"RdYlBu")
display.brewer.pal(11,"RdYlGn")
```


## Rotate in three dimensions
The app allow user to choose the value for longitude, latitude and angle. As we know, the earth is a three-axial ellipsoid.

- Longitude will make the model roate around polar axis, the value range from -180 to 180
- Latitude will make the model rotate around one diameter on equator plane, the value range from -90 to 90, at 90 the model will show teh north pole facing and -90 the south pole
- Angle will make the model rotate around another diameter on equator plane, the value range from -180 to 180

## Example
We will show a example of the model given the color palette and three values.  
Palette=Pink Yellow Green, Longtitude=90,Latitude=45,Angle=0
```{r,echo = FALSE}
suppressPackageStartupMessages(library(ggplot2))

world_map<-map_data("world")
world_map$region<-factor(world_map$region)
Pal="PiYG"
Lat=45
Long=90
Angle=0
g<- ggplot(data=world_map) + geom_polygon(aes(x=long,y=lat,group=group,fill=as.numeric(region)),
                             colour="grey",size=.25)
      g<- g + coord_map("ortho",orientation=c(Lat,Long,Angle))
      g<- g + scale_y_continuous(breaks=(-6:6)*15) 
      g<- g + scale_x_continuous(breaks=(-12:12)*15)
      g<- g + scale_fill_distiller(name="Region ID",palette=Pal)
      g<- g + theme(
          panel.background=element_rect(fill="white",colour=NA),
          panel.grid.major = element_line(colour = "grey60",size=.25),
          panel.grid.minor = element_line(colour = "grey60",size=.25),
          axis.text=element_blank(),
          axis.title=element_blank(),
          axis.ticks=element_blank())
      g
```
