---
title: "Week1 Quiz"
author: "Shawvin"
date: "7/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Quiz 1
### Quiz, 10 questions

1.Question 1

Consider the data set given below 
```{r}
x <- c(0.18, -1.54, 0.42, 0.95)
print(paste("x <-",list(x)))
```

And weights given by
```{r}
w <- c(2, 1, 3, 1)
print(paste("w <-",list(w)))
```

Give the value of μ that minimizes the least squares equation
$\sum_{i=1}^{n}w_{i}(x_{i}-u)^{2}$

Answer:
```{r}
result<-sum(x*w)/sum(w)
result
```


2.Question 2

Consider the following data set
```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
print(paste("x <-",list(x)))
print(paste("y <-",list(y)))
```
Fit the regression through the origin and get the slope treating y as the outcome and x as the regressor. (Hint, do not center the data since we want regression through the origin, not through the means of the data.)



3. Question 3

Do data(mtcars)from the datasets package and fit the regression model with mpg as the outcome and weight as the predictor. Give the slope coefficient.

```{r}
coeff<-cor(mtcars$wt,mtcars$mpg)
coeff
```
